---
title: "DATA FEST 2024"
author: "Haojie Liu"
date: "2024-04-27"
output: pdf_document
---

```{r}
library(tidyverse)
library(forecast)
```

```{r}

checkpoints_eoc <- read.csv("checkpoints_eoc.csv")
checkpoints_pulse <- read.csv("checkpoints_pulse.csv")
items <- read.csv("items.csv")
media_views <- read.csv("media_views.csv")
page_views <- read.csv("page_views.csv")
responses <- read.csv("responses.csv")

```

```{r}
checkpoints_eoc %>% 
  head()

items
```


```{r}
 # check if students have more than one institution


checkpoints_eoc %>% 
  left_join(checkpoints_pulse, by= join_by(student_id)) %>% 
  group_by(student_id) %>% 
  select(student_id, institution_id) %>% 
  distinct() %>% 
  na.omit() %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
  
```



```{r}

checkpoints_eoc %>% 
  left_join(checkpoints_pulse, by= join_by(student_id)) %>%
  mutate(institution_id = ifelse(is.na(institution_id), "self", institution_id)) %>% 
  group_by(institution_id) %>% 
  summarize(count = n(), 
            mean_score = mean(EOC, na.rm = TRUE)) %>% 
  arrange(desc(mean_score))
  
```


```{r}
checkpoints_eoc %>% 
  left_join(items, by = join_by(class_id))

items

checkpoints_eoc

media_views %>% 
  select(student_id, book)

book_collect <- page_views %>% 
  select(class_id, book, release) %>% 
  group_by(class_id) %>% 
  unique()


temp <- checkpoints_eoc %>% 
  group_by(class_id, chapter_number) %>% 
  summarize(mean_score = mean(EOC, na.rm = TRUE)) %>% 
  left_join(book_collect, join_by(class_id))


temp %>% 
  filter(book == "College / Statistics and Data Science (ABC)") %>% 
  filter(class_id == "20bd524c-bb2d-4b74-a419-929475b91d94")


temp %>% 
  filter(book == "College / Statistics and Data Science (ABC)") %>% 
  ggplot(aes(
    x = chapter_number,
    y = mean_score,
    color = class_id
  )) + geom_line() + theme(legend.position="none")

lm1 <- with(temp %>% filter(book == "College / Statistics and Data Science (ABC)"), lm(`mean_score` ~ chapter_number))

summary(lm1)
plot(lm1)


```



```{r}
page_views %>% 
  filter(!is.na(chapter_number)) %>% 
  mutate(engaged = log(abs(engaged)))%>% 
  mutate(engaged = ifelse(engaged == -Inf, 0.0001, engaged)) %>% 
  group_by(class_id, chapter_number) %>% 
  summarize(mean_time = mean(engaged, na.rm = TRUE)) %>% 
  left_join(book_collect, join_by(class_id)) %>% 
  filter(book == "College / Statistics and Data Science (ABC)") %>% 
  ggplot(aes(
    x = chapter_number,
    y = mean_time,
    color = class_id
  )) + geom_line() + theme(legend.position="none")


temp <- page_views %>% 
  filter(!is.na(chapter_number)) %>% 
  mutate(engaged = log(abs(engaged)))%>% 
  mutate(engaged = ifelse(engaged == -Inf, 0.0001, engaged)) %>% 
  group_by(class_id, chapter_number) %>% 
  summarize(mean_time = mean(engaged, na.rm = TRUE)) %>% 
  left_join(book_collect, join_by(class_id))

summary(lm(mean_time ~ chapter_number, data = temp))


```


```{r}
page_views %>% 
  filter(!is.na(chapter_number)) %>% 
  group_by(class_id, chapter_number)
page_views

for(i in 1:12){
  a <- page_views %>% filter(book == "College / Statistics and Data Science (ABC)", chapter_number == i) %>% select(engaged)
  a <- log(abs(a))

  hist(unlist(a), breaks = 100)
}


page_views %>% 
  filter(!is.na(chapter_number)) %>% 
  group_by(class_id, chapter_number) %>% 
  summarize(a = log(abs(engaged))) %>% 
  na.omit() %>% 
  filter(chapter_number == 2) %>% 
  mutate(a = ifelse(a == -Inf, 0.0001, a)) %>% 
  summarize(mean(a))
  
  
  summarize(mean_time = mean(log(abs(engaged)), na.rm = TRUE))
  
  
  

```




objective: 
  Find variables that will somehow influence "engaged", which represent how long a student can be stucked on a single chapter page
  

variables:


response var:

EOC
engaged

  idle_brief
  idle_long


```{r}
table(page_views$chapter_number, page_views$section_number, page_views$was_complete)


temp <- page_views %>% 
  group_by(book, chapter_number, section_number) %>% 
  summarise(view_n = sum(was_complete, na.rm = TRUE)) %>% 
  na.omit() %>% 
  arrange(desc(view_n))


plot(temp$chapter_number, temp$view_n)


temp <- page_views %>% 
  filter(book=="College / Advanced Statistics and Data Science (ABCD)") %>% 
  group_by(student_id,chapter_number, section_number) %>% 
  summarize(tried_again_clicks = mean(tried_again_clicks)) %>% 
  ungroup() %>% 
  group_by(chapter_number, section_number) %>% 
  summarise(view_n = mean(tried_again_clicks, na.rm = TRUE)) %>% 
  na.omit() %>% 
  arrange(desc(view_n))


plot(temp$chapter_number, temp$view_n)

page_views

```



```{r}

temp <- page_views %>% 
  group_by(book, student_id,chapter_number, section_number) %>% 
  summarize(tried_again_clicks = mean(tried_again_clicks)) %>% 
  ungroup() %>% 
  group_by(chapter_number, section_number) %>% 
  summarise(view_n = mean(tried_again_clicks, na.rm = TRUE)) %>% 
  na.omit() %>% 
  arrange(desc(view_n))

temp


plot(temp$chapter_number, temp$view_n)

page_views


```



```{r}
responses %>% 
  group_by(student_id) %>% 
  summarise(m = mean(attempt),
            maxi = max(attempt)) %>% 
  arrange(desc(m))


responses %>% 
  group_by(item_type) %>% 
  summarise(m = mean(attempt),
            maxi = max(attempt)) %>% 
  arrange(desc(m))

temp2 <- responses %>% 
  filter(item_type != "code") %>% 
  group_by(student_id, chapter_number, section_number, item_id) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))


responses %>% 
  select(student_id, chapter_number, section_number, item_id, points_earned)

responses
```

```{r}
responses
```


on question
correct

is the student earned this point because that he try many many attempts
if yes, can we use a method to avoid this happen, for example generate a similar question that can never repeat

non-correct

find question that cause many student lose point,
see if the question are misleading or concept are too hard(question bias)
if yes, can we improve
if not is it because of the topic is too hard or textbook explain the concept too complex



```{r}
responses %>% 
  filter(item_type != "code") %>% 
  filter(points_possible == points_earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(chapter_number,
            section_number,
            attempt = n())

responses %>% 
  filter(item_type != "code")

responses %>% 
  filter(item_type != "code") %>% 
  filter(points_possible > points_earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(chapter_number,
            section_number,
            attempt = n())


new_responses <- responses %>% 
  filter(item_type != "code") %>% 
  group_by(student_id, item_id) %>% 
  summarize(earned = any(points_possible == points_earned)) %>% 
  right_join(responses %>% filter(item_type != "code"), by = join_by(student_id, item_id))



new_responses %>% 
  ungroup() %>% 
  filter(earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(attempt = n()) %>% 
  ggplot(
    aes(
      x = attempt
    )
  ) + geom_histogram()

new_responses %>% 
  ungroup() %>% 
  filter(earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(attempt = n()) %>% 
  arrange(desc(attempt))


responses %>% 
  filter(item_type == "code") %>% 
  filter(points_possible == points_earned) %>% 
  ggplot(
    aes(
      x = attempt
    )
  ) + geom_histogram()

responses %>% 
  filter(item_type == "code") %>% 
  filter(points_possible > points_earned) %>% 
  ggplot(
    aes(
      x = attempt
    )
  ) + geom_histogram()


responses %>% 
  filter(item_type == "code") %>% 
  group_by(student_id, item_id) %>% 
  summarize(earned = any(points_possible == points_earned)) %>% 
  right_join(responses %>% filter(item_type == "code"), by = join_by(student_id, item_id)) %>% 
  ungroup() %>% 
  filter(earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(attempt = n()) %>% 
  arrange(desc(attempt)) %>% 
  mutate(range = case_when(
    attempt >= 15 ~ "15 or more",
    attempt >= 10 ~ "10 to 15",
    attempt >= 5 ~ "5 to 10",
    attempt >= 0 ~ "less than 5"
  )) 


responses %>% 
  filter(item_type != "code") %>% 
  group_by(student_id, item_id) %>% 
  summarize(earned = any(points_possible == points_earned)) %>% 
  right_join(responses %>% filter(item_type != "code"), by = join_by(student_id, item_id)) %>% 
  ungroup() %>% 
  filter(earned) %>% 
  group_by(student_id, item_id) %>% 
  summarize(attempt = n()) %>% 
  arrange(desc(attempt)) %>% 
  mutate(range = case_when(
    attempt >= 15 ~ "15 or more",
    attempt >= 10 ~ "10 to 15",
    attempt >= 5 ~ "5 to 10",
    attempt >= 0 ~ "less than 5"
  )) %>% 
  write.csv("non_code_attempt.csv")


```
```{r}
new_responses %>%
  ungroup() %>%
  filter(earned) %>%
  group_by(student_id,item_id) %>%
  summarize(attempt = n()) %>%
  ggplot(aes(x="", y=attempt, fill=factor(attempt))) + geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") + theme_void() + labs(fill = "attempts")
```



```{r}
temp <- responses %>% 
  group_by(student_id, item_id) %>% 
  summarize(earned = any(points_possible == points_earned)) %>% 
  right_join(responses, by = join_by(student_id, item_id))



temp %>%
  filter(!earned) %>% 
  ggplot(
    aes(
      x = chapter_number,
      fill = book
    )
  ) + geom_histogram(position = "dodge") + theme(legend.position="top")



temp <- checkpoints_eoc %>% 
  group_by(book, chapter_number) %>% 
  summarize(chapter_number,
            EOC = mean(EOC, na.rm = TRUE),
            book) %>% 
  unique()

temp %>% 
  group_by(book) %>% 
  subset(book == "College / Advanced Statistics and Data Science (ABCD)") %>% 
  tsdisplay()

temp %>% 
  group_by(book) %>% 
  subset(book == "College / Statistics and Data Science (ABC)") %>% 
  tsdisplay()

temp %>% 
  group_by(book) %>% 
  subset(book == "High School / Advanced Statistics and Data Science I (ABC)") %>% 
  tsdisplay()
```


```{r}
checkpoints_eoc %>% 
  group_by(chapter_number, book) %>% 
  summarize(chapter_number,
            EOC = mean(EOC, na.rm = TRUE),
            book) %>% 
  ggplot(
    aes(
      x = chapter_number,
      y = EOC,
      color = book
    )
  ) + geom_line(size=2) + theme_classic() + theme(legend.position="left", legend.key.size =  unit(0.5,"line"))

ggsave("line.png")

```


```{r}

responses

```



engage vs EOS

```{r warning=FALSE}
checkpoints_eoc

x <- page_views %>% 
  group_by(book, student_id, chapter_number) %>% 
  summarise(total_engaged = sum(engaged, na.rm = TRUE)) %>% 
  right_join(checkpoints_eoc, by = join_by(student_id, chapter_number, book)) 

x %>% 
  ungroup() %>% 
  group_by(chapter_number) %>% 
  summarize(mean_total_engaged = mean(total_engaged, na.rm = TRUE),
            mean_EOC = mean(EOC, na.rm = TRUE))%>% 
  ggplot(
    aes(
      x = mean_total_engaged,
      y = mean_EOC
    )
  ) + geom_line()


for (i in 1:16) {
  p <- x %>% 
    filter(chapter_number == i) %>% 
    ggplot(
    aes(
      x = total_engaged,
      y = EOC
    )
  ) + geom_point()
  
  print(p)
  
  a <- summary(lm(EOC~ total_engaged, data = x %>% filter(chapter_number == i)))
  print(a)
}
```

```{r}

checkpoints_eoc <- checkpoints_eoc %>%
  group_by(class_id, chapter_number) %>%
  mutate(
    # Calculate quantiles within each group
    quantile = ntile(EOC, 4),  # Divides into quartiles

    # Assign status based on quantiles
    status = case_when(
      EOC > 0.7 ~ "On Progress",
      quantile == 1 ~ "Fall Behind",
      quantile == 2 ~ "Risky",
      quantile >= 3 ~ "On Progress"
    )
  )
checkpoints_eoc %>% 
  write.csv("checkpoints_eoc_status.csv")

```


```{r}
page_views
```

student id
attempt

Book and Chapter

Engaged idle_brief, idle_long, off_page_brief, off_page_long (sum up by chapter)

```{r}
new_page_views <- page_views %>% 
  group_by(student_id, book, chapter_number) %>% 
  summarize(total_engaged = sum(engaged, na.rm = TRUE),
            total_idel_brief = sum(idle_brief, na.rm = TRUE),
            total_idle_long = sum(idle_long, na.rm = TRUE),
            total_off_page_brief = sum(off_page_brief, na.rm = TRUE),
            total_off_page_long = sum(off_page_long, na.rm = TRUE)
            )

new_page_views
```

```{r}

new_tried_again_clicks <- page_views %>% 
  group_by(student_id, book, chapter_number, section_number) %>% 
  summarize(x = max(tried_again_clicks)) %>% 
  ungroup() %>% 
  group_by(student_id, book, chapter_number) %>% 
  summarize(total_tried_again_clicks = sum(x))

new_was_complete <- page_views %>% 
  mutate(was_complete_true = ifelse(was_complete, 1, 0),
         was_complete_false = ifelse(!was_complete, 1, 0)) %>% 
  group_by(student_id, book, chapter_number) %>% 
  summarize(was_complete_true = sum(was_complete_true), 
            was_complete_false = sum(was_complete_false))
  

```


tried_again_clicks (max by section, sum by chapter)
was_complete(dummy then sum())

EOC from lag1 chapter()


```{r}
temp_eoc <- checkpoints_eoc %>%
  arrange(student_id, chapter_number) %>%  # Make sure data is ordered correctly
  group_by(student_id) %>%
  mutate(lag1_eoc = lag(EOC, 1),
         lag2_eoc = lag(EOC, 2),
         lag3_eoc = lag(EOC, 3))
```

```{r}

new_responses <- responses %>%
  group_by(student_id, book, chapter_number) %>%
  summarise(total_attempt = sum(attempt, na.rm = TRUE))

```


```{r}
df_full <- new_page_views %>% 
  inner_join(new_tried_again_clicks, by = join_by(student_id, chapter_number)) %>% 
  inner_join(new_was_complete, by = join_by(student_id, chapter_number)) %>% 
  inner_join(new_responses,  by = join_by(student_id,book , chapter_number)) %>% 
  right_join(temp_eoc, by = join_by(student_id, book, chapter_number)) %>% 
  select(-c(quantile, book.x, book.y, EOC, class_id))


df_full <- df_full %>% 
  mutate(book = as.factor(book),
         status = as.factor(status))

write.csv(df_full, "df_full.csv")
```


```{r}
df_full
```



```{r}
responses %>% 
  filter(item_type != "code")
```









